\documentclass[11pt,twoside]{article}

%\documentclass{sig-alternate}

\usepackage{jeffe,graphicx,hyperref}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{marvosym}
\usepackage{cite}
\usepackage{verbatim}

%-----------------------------------------------------------------------
%  These packages are purely cosmetic.
%  Just comment them out if they don't work for you.
%-----------------------------------------------------------------------
\usepackage{microtype}
\usepackage[charter]{mathdesign}
\def\sfdefault{fvs}
\def\ttdefault{fvm}
\usepackage[mathcal]{euscript}
\usepackage{stmaryrd}


%-----------------------------------------------------------------------
%  Local definitions
%-----------------------------------------------------------------------
\def\arcto{\mathord\shortrightarrow}
\def\arc#1#2{#1\arcto#2}
\def\cra#1#2{#1\mathord\shortleftarrow#2}
\def\fence#1#2{#1\mathord\shortuparrow#2}
\def\ecnef#1#2{#1\mathord\shortdownarrow#2}
\def\head{\mathit{head}}
\def\tail{\mathit{tail}}
\def\lsh{\mathit{left}}
\def\rsh{\mathit{right}}
\def\rev{\mathit{rev}}
\def\Z{\mathbb{Z}}
\def\Real{\mathbb{R}}
\def\Q{\mathbb{Q}}

% needs marvosym package:
\def\snip{\mathbin{\raisebox{0.15ex}{\rotatebox[origin=c]{60}{\Rightscissors}\!}}}
\let\unlhd\trianglelefteq	% fix mathdesign font bug

\let\cycle\gamma
\let\path p
\let\primalarc\alpha
\def\dualarc{p}

\def\Sigmabar{\overline{\smash{\Sigma}\vphantom{t}}}
\def\SIGMABAR{\boldsymbol{\overline{\smash{\Sigma}\vphantom{x}}}}
\def\Gbar{\overline{\smash{G}\vphantom{t}}}
\def\Vbar{\overline{\smash{V}\vphantom{t}}}
\def\Ebar{\overline{\smash{E}\vphantom{t}}}
\def\bbar{\overline{\smash{b}\vphantom{t}}}
\def\nbar{\overline{n}}
\def\gbar{\overline{g}}
\def\wbar{\overline{w}}
\def\sigmabar{\overline{\sigma}}
\def\cyclebar{\overline{\cycle}}
\def\chibar{\overline{\chi}}

\def\fakeparagraph#1{\par\medskip\noindent\textbf{#1}}


\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\pagestyle{myheadings}
\markboth{Minimum cuts in surface graphs}
		 {Erin W. Chambers, Jeff Erickson, Kyle Fox, and Amir Nayyeri}

\begin{titlepage}

\title{Minimum cuts in surface graphs\footnote{
Portions of this work done by different subsets of the authors appeared in~\cite{cen-mcshc-09},~\cite{en-mcsnc-11}, and~\cite{efn-gmcse-12}.
}}

\author{
  Erin W. Chambers%
  \thanks{Department of Computer Science and Mathematics, Saint Louis
  University;
  \url{echambe5@slu.edu}.  Supported in part by NSF grants CCF 1054779 and DMS-0528086.
  Portions of this work were done while this author was a student at the University of Illinois at Urbana-Champaign.}
  \and
  Jeff Erickson%
  \thanks{Department of Computer Science,
  University of Illinois, Urbana-Champaign; \url{jeffe@illinois.edu}.
  Supported in part by NSF grants CCF 09-15519 and DMS-0528086.
  }
  \and
  Kyle Fox%
  \thanks{Department of Computer Science,
      University of Illinois, Urbana-Champaign;
      \url{kylefox2@illinois.edu}.
      Supported in part by
      the Department of Energy Office
      of Science Graduate Fellowship Program (DOE SCGF),
      made possible in part by the American Recovery and
      Reinvestment Act of 2009, administered by ORISE-ORAU
      under contract no. DE-AC05-06OR23100.}
  \and
  Amir Nayyeri%
  \thanks{Department of Computer Science,
      Carnegie Mellon University,
      \url{amirn@cs.cmu.edu}. Supported in part by NSF grants
      CCF 1065106, CCF 09-15519, and DMS-0528086. Portions of this work were done while this author was a student 
      at the University of Illinois at Urbana-Champaign.}
      }

\DRAFT

\maketitle
\begin{abstract}
Let $G$ be an edge-weighted directed graph with $n$ vertices embedded on an orientable surface of genus $g$.
We describe algorithms to efficiently compute minimum $s,t$-cuts and global minimum cuts of~$G$.
\note{TODO(kylejfox): Rewrite abstract.}
\end{abstract}

\noindent

\thispagestyle{empty}
\setcounter{page}{0}
\end{titlepage}


%---------------------------------------
\input{intro}
\input{prelims}
\input{characterizing}
\input{crossing}
\input{homcover}
\input{hardness}
\input{global}
\input{conclusions}



% Begin bbl nonsense
\bibliographystyle{abbrv}
\bibliography{topology,data-structures,optimization,other}


\end{document}

