\def\burl#1{$\langle$\url{#1}$\rangle$}
\begin{thebibliography}{10}
\itemsep 1pt plus 1pt

\bibitem{amo-nftaa-93}
R.~K. Ahuja, T.~L. Magnanti, and J.~Orlin.
\newblock \emph{Network Flows: Theory, Algorithms, and Applications}.
\newblock Prentice Hall, 1993.

\bibitem{b-epnfc-08}
G.~Borradaile.
\newblock \emph{Exploiting Planarity for Network Flow and Connectivity
  Problems}.
\newblock Ph.D. thesis, Brown University, May 2008.
\newblock
  \burl{http://www.cs.brown.edu/research/pubs/theses/phd/2008/glencora.pdf}.

\bibitem{bdt-ptass-08}
G.~Borradaile, E.~D. Demaine, and S.~Tazari.
\newblock Polynomial-time approximation schemes for subset-connectivity
  problems in bounded-genus graphs.
\newblock \emph{Proc. 26th Int. Symp. Theoretical Aspects Comput. Sci.},
  171--182, 2009. Dagstuhl Seminar Proceedings.
\newblock \burl{http://drops.dagstuhl.de/opus/volltexte/2009/1835/}.

\bibitem{bkk-ptass-07}
G.~Borradaile, C.~Kenyon-Mathieu, and P.~N. Klein.
\newblock A polynomial-time approximation scheme for {Steiner} tree in planar
  graphs.
\newblock \emph{Proc. 18th Ann. ACM-SIAM Symp. Discrete Algorithms},
  1285--1294, 2007.

\bibitem{bkk-stpg-07}
G.~Borradaile, C.~Kenyon-Mathieu, and P.~N. Klein.
\newblock Steiner tree in planar graphs: {An} {$O(n \log n)$} approximation
  scheme with singly-exponential dependence on epsilon.
\newblock \emph{Proc. 10th Ann. Workshop on Algorithms and Data Structures},
  275--286, 2007.

\bibitem{bk-tamfd-06}
G.~Borradaile and P.~Klein.
\newblock An {$O(n \log n)$}-time algorithm for maximum $st$-flow in a directed
  planar graph.
\newblock \emph{Proc. 17th Ann. ACM-SIAM Symp. Discrete Algorithms}, 524--533,
  2006.

\bibitem{bk-amfdp-09}
G.~Borradaile and P.~Klein.
\newblock An {$O(n\log n)$} algorithm for maximum $st$-flow in a directed
  planar graph.
\newblock \emph{J. ACM} $\!\!$, to appear, 2009.
\newblock
  \burl{http://www.math.uwaterloo.ca/~glencora/downloads/maxflow-full.pdf}.

\bibitem{c-mdpg-06}
S.~Cabello.
\newblock Many distances in planar graphs.
\newblock \emph{Proc. 17th Ann. ACM-SIAM Symp. Discrete Algorithms},
  1213--1220, 2006.

\bibitem{cc-msspg-07}
S.~Cabello and E.~W. Chambers.
\newblock Multiple source shortest paths in a genus $g$ graph.
\newblock \emph{Proc. 18th Ann. ACM-SIAM Symp. Discrete Algorithms}, 89--97,
  2007.

\bibitem{cdem-fotc-08}
S.~Cabello, M.~DeVos, J.~Erickson, and B.~Mohar.
\newblock Finding one tight cycle.
\newblock \emph{Proc. 19th Ann. ACM-SIAM Symp. Discrete Algorithms}, 527--531,
  2008.

\bibitem{cm-fsnsn-07}
S.~Cabello and B.~Mohar.
\newblock Finding shortest non-separating and non-contractible cycles for
  topologically embedded graphs.
\newblock \emph{Discrete Comput. Geom.} 37:213--235, 2007.

\bibitem{ccelw-scsih-06}
E.~W. Chambers, {\'E}.~{Colin de Verdi\`{e}re}, J.~Erickson, F.~Lazarus, and
  K.~Whittlesey.
\newblock Splitting (complicated) surfaces is hard.
\newblock \emph{Proc. 22nd Ann. ACM Symp. Comput. Geom.}, 421--429, 2006.

\bibitem{ccelw-scsih-08}
E.~W. Chambers, {\'E}.~{Colin de Verdi{\`e}re}, J.~Erickson, F.~Lazarus, and
  K.~Whittlesey.
\newblock Splitting (complicated) surfaces is hard.
\newblock \emph{Comput. Geom. Theory Appl.} 41(1--2):94--110, 2008.

\bibitem{cen-hfcc-09}
E.~W. Chambers, J.~Erickson, and A.~Nayyeri.
\newblock Homology flows, cohomology cuts.
\newblock \emph{Proc. 41st Ann. ACM Symp. Theory Comput.}, 2009.

\bibitem{cf-qhc2-07}
C.~Chen and D.~Freedman.
\newblock Quantifying homology classes {II}: {L}ocalization and stability.
\newblock Preprint, 2007.
\newblock \burl{http://arxiv.org/abs/0709.2512v2}.

\bibitem{cf-qhc-08}
C.~Chen and D.~Freedman.
\newblock Quantifying homology classes.
\newblock \emph{Proc. 25th Ann. Symp. Theoretical Aspects Comp. Sci.},
  169--180, 2008. Dagstuhl Seminar Proceedings.
\newblock \burl{http://drops.dagstuhl.de/opus/volltexte/2008/1343/}.

\bibitem{c-rcds-03}
{\'E}.~{Colin de Verdi{\`e}re}.
\newblock \emph{Raccourcissement de courbes et d{\'e}composition de surfaces
  [Shortening of Curves and Decomposition of Surfaces]}.
\newblock Ph.D. thesis, University of Paris 7, Dec. 2003.
\newblock \burl{http://www.di.ens.fr/~colin/textes/these.html}.

\bibitem{ce-tspcs-06}
{\'E}.~{Colin de Verdi{\`e}re} and J.~Erickson.
\newblock Tightening non-simple paths and cycles on surfaces.
\newblock \emph{Proc. 17th Ann. ACM-SIAM Symp. Discrete Algorithms}, 192--201,
  2006.

\bibitem{cl-opdsh-03}
{\'E}.~{Colin de Verdi{\`e}re} and F.~Lazarus.
\newblock Optimal pants decompositions and shortest homotopic cycles on an
  orientable surface.
\newblock \emph{Proc. 11th Sympos. Graph Drawing}, 478--490, 2003. Lecture
  Notes Comput. Sci. 2912.

\bibitem{cl-oslos-05}
{\'E}.~{Colin de Verdi{\`e}re} and F.~Lazarus.
\newblock Optimal system of loops on an orientable surface.
\newblock \emph{Discrete Comput. Geom.} 33(3):507--534, 2005.

\bibitem{cl-opdsh-07}
{\'E}.~{Colin de Verdi{\`e}re} and F.~Lazarus.
\newblock Optimal pants decompositions and shortest homotopic cycles on an
  orientable surface.
\newblock \emph{J. ACM} 54(4), 2007.

\bibitem{dhm-aacd-07}
E.~D. Demaine, M.~Hajiaghayi, and B.~Mohar.
\newblock Approximation algorithms via contraction decomposition.
\newblock \emph{Proc. 18th Ann. ACM-SIAM Symp. Discrete Algorithms ACM-SIAM
  symposium on Discrete algorithms}, 278--287, 2007.

\bibitem{dls-chtl-07}
T.~K. Dey, K.~Li, and J.~Sun.
\newblock On computing handle and tunnel loops.
\newblock \emph{IEEE Proc. Int. Conf. Cyberworlds}, 357--366, 2007.

\bibitem{dlsc-cgaht-08}
T.~K. Dey, K.~Li, J.~Sun, and D.~Cohen-Steiner.
\newblock Computing geometry-aware handle and tunnel loops in {3D} models.
\newblock \emph{ACM Trans. Graphics} 27(3):1--9, 2008.
\newblock Proc. SIGGRAPH 2008.

\bibitem{e-sipgr-99}
D.~Eppstein.
\newblock Subgraph isomorphism in planar graphs and related problems.
\newblock \emph{J. Graph Algorithms and Applications} 3(3):1--27, 1999.

\bibitem{e-dtmcg-00}
D.~Eppstein.
\newblock Diameter and treewidth in minor-closed graph families.
\newblock \emph{Algorithmica} 27:275--291, 2000.

\bibitem{eh-ocsd-04}
J.~Erickson and S.~Har-Peled.
\newblock Optimally cutting a surface into a disk.
\newblock \emph{Discrete Comput. Geom.} 31(1):37--59, 2004.

\bibitem{ew-gohhg-05}
J.~Erickson and K.~Whittlesey.
\newblock Greedy optimal homotopy and homology generators.
\newblock \emph{Proc. 16th Ann. ACM-SIAM Symp. Discrete Algorithms},
  1038--1046, 2005.

\bibitem{ew-csec-08}
J.~Erickson and P.~Worah.
\newblock Computing the shortest essential cycle.
\newblock Preprint, November 2008.
\newblock \burl{http://www.cs.uiuc.edu/~jeffe/pubs/essential.html}.

\bibitem{fr-pgnwe-06}
J.~Fakcharoenphol and S.~Rao.
\newblock Planar graphs, negative weight edges, shortest paths, and near linear
  time.
\newblock \emph{J. Comput. Syst. Sci.} 72(5):868--889, 2006.

\bibitem{ff-mfn-56}
L.~R. Ford and D.~R. Fulkerson.
\newblock Maximal flow through a network.
\newblock \emph{Canadian J. Math.} 8(399--404), 1956.

\bibitem{f-faspp-87}
G.~N. Frederickson.
\newblock Fast algorithms for shortest paths in planar graphs with
  applications.
\newblock \emph{SIAM J. Comput.} 16(6):1004--1004, 1987.

\bibitem{fgm-fmggi-88}
M.~L. Furst, J.~L. Gross, and L.~A. McGeoch.
\newblock Finding a maximum-genus graph imbedding.
\newblock \emph{J. Assoc. Comput. Mach.} 35(3):523--534, 1988.

\bibitem{gr-bfdb-98}
A.~V. Goldberg and S.~Rao.
\newblock Beyond the flow decomposition barrier.
\newblock \emph{J. ACM} 45(5):783--797, 1998.

\bibitem{gt-namfp-88}
A.~V. Goldberg and R.~E. Tarjan.
\newblock A new approach to the maximum-flow problem.
\newblock \emph{J. Assoc. Comput. Mach.} 35(4):921--940, 1988.

\bibitem{g-itegd-00}
M.~Grohe.
\newblock Isomorphism testing for embeddable graphs through definability.
\newblock \emph{Proc. 32nd ACM Symp. Theory Comput.}, 63--72, 2000.

\bibitem{gt-tgt-01}
J.~L. Gross and T.~W. Tucker.
\newblock \emph{Topological graph theory}.
\newblock Dover Publications, 2001.

\bibitem{gw-tnr-01}
I.~Guskov and Z.~Wood.
\newblock Topological noise removal.
\newblock \emph{Proc. Graphics Interface}, 19--26, 2001.

\bibitem{hr-fmern-55}
T.~E. Harris and F.~S. Ross.
\newblock Fundamentals of a method for evaluating rail net capacities.
\newblock Tech. rep., The RAND Corporation, Santa Monica, California, October
  24 1955.
\newblock Cited in \cite{s-hco-05}.

\bibitem{hj-oamfu-85}
R.~Hassin and D.~B. Johnson.
\newblock An {$O(n\log^2 n)$} algorithm for maximum flow in undirected planar
  networks.
\newblock \emph{SIAM J. Comput.} 14(3):612--624, 1985.

\bibitem{h-at-01}
A.~Hatcher.
\newblock \emph{Algebraic Topology}.
\newblock Cambridge University Press, 2001.
\newblock \burl{http://www.math.cornell.edu/~hatcher/}.

\bibitem{hkrs-fspap-97}
M.~R. Henzinger, P.~Klein, S.~Rao, and S.~Subramanian.
\newblock Faster shortest-path algorithms for planar graphs.
\newblock \emph{J. Comput. Syst. Sci.} 55(1):3--23, 1997.

\bibitem{hw-mstfkc-07}
J.~M. Hochstein and K.~Weihe.
\newblock Maximum $s$-$t$-flow with $k$ crossings in {$O(k^3n\log n)$} time.
\newblock \emph{Proc. 18th Ann. ACM-SIAM Symp. Discrete Algorithms}, 843--847,
  2007.

\bibitem{hw-ltaip-74}
J.~E. Hopcroft and J.~K. Wong.
\newblock Linear time algorithm for isomorphism of planar graphs (preliminary
  report).
\newblock \emph{Proc. 6th ACM Symp. Theory Comput.}, 172--184, 1974.

\bibitem{ii-espap-90}
H.~Imai and K.~Iwano.
\newblock Efficient sequential and parallel algorithms for planar minimum cost
  flow.
\newblock \emph{Proc. SIGAL Int. Symp. Algorithms}, 21--30, 1990. Lecture Notes
  Comput. Sci. 450, Springer-Verlag.

\bibitem{is-pebgg-07}
P.~Indyk and A.~Sidiropoulos.
\newblock Probabilistic embeddings of bounded genus graphs into planar graphs.
\newblock \emph{Proc. 23rd Ann. ACM Symp. Comput. Geom.}, 204--209, 2007.

\bibitem{is-mfpn-79}
A.~Itai and Y.~Shiloach.
\newblock Maximum flow in planar networks.
\newblock \emph{SIAM J. Comput.} 8:135--150, 1979.

\bibitem{jk-mcdpn-92}
L.~Janiga and V.~Koubek.
\newblock Minimum cut in directed planar networks.
\newblock \emph{Kybernetika} 28(1):37--49, 1992.

\bibitem{kr-ccnlt-07}
K.~Kawarabayashi and B.~Reed.
\newblock Computing crossing number in linear time.
\newblock \emph{Proc. 39th Ann ACM Symp. Theory Comput.}, 382--390, 2007.

\bibitem{k-msspp-05}
P.~Klein.
\newblock Multiple-source shortest paths in planar graphs.
\newblock \emph{Proc. 16th Ann. ACM-SIAM Symp. Discrete Algorithms}, 146--155,
  2005.

\bibitem{kmw-spdpg-09}
P.~Klein, S.~Mozes, and O.~Weimann.
\newblock Shortest paths in directed planar graphs with negative lengths: {A}
  linear-space {$O(n\log^2 n)$}-time algorithm.
\newblock \emph{Proc. 20th Ann. ACM-SIAM Symp. Discrete Algorithms}, 236--245,
  2009.

\bibitem{k-csnco-06}
M.~Kutz.
\newblock Computing shortest non-trivial cycles on orientable surfaces of
  bounded genus in almost linear time.
\newblock \emph{Proc. 22nd Ann. ACM Symp. Comput. Geom.}, 430--438, 2006.

\bibitem{lrt-gnd-79}
R.~J. Lipton, D.~J. Rose, and R.~E. Tarjan.
\newblock Generalized nested dissection.
\newblock \emph{SIAM J. Numer. Anal.} 16:346--358, 1979.

\bibitem{m-tltam-04}
M.~Mare{\v s}.
\newblock Two linear time algorithms for {MST} on minor closed graph classes.
\newblock \emph{Archivum Mathematicum} 40(3):315--320, 2004.

\bibitem{mrr-edofm-03}
S.~T. McCormick, M.~R. Rao, and G.~Rinaldi.
\newblock Easy and difficult objective functions for max cut.
\newblock \emph{Math. Program., Ser. B} 94:459--466, 2003.

\balancecolumns
\itemsep 1.5pt

\bibitem{m-itgbg-80}
G.~L. Miller.
\newblock Isomorphism testing for graphs of bounded genus.
\newblock \emph{Proc. 12th ACM Symp. Theory Comput.}, 225--235, 1980.

\bibitem{mn-fpgms-95}
G.~L. Miller and J.~Naor.
\newblock Flow in planar graphs with multiple sources and sinks.
\newblock \emph{SIAM J. Comput.} 24(5):1002--10017, 1995.

\bibitem{mt-gos-01}
B.~Mohar and C.~Thomassen.
\newblock \emph{Graphs on Surfaces}.
\newblock Johns Hopkins University Press, 2001.

\bibitem{mvv-memi-87}
K.~Mulmuley, U.~Vazirani, and V.~Vazirani.
\newblock Matching is as easy as matrix inversion.
\newblock \emph{Combinatorica} 7:105--113, 1987.

\bibitem{pr-fepss-93}
V.~Y. Pan and J.~H. Reif.
\newblock Fast and efficient parallel solution of sparse linear systems.
\newblock \emph{SIAM J. Comput.} 22(6):1227--1250, 1993.

\bibitem{p-omst-99}
D.~Pe'er.
\newblock On minimum spanning trees.
\newblock Master's thesis, Hebrew University, 1998.
\newblock \burl{http://www.math.ias.edu/~avi/STUDENTS/dpthesis.pdf}.

\bibitem{r-mstcp-83}
J.~Reif.
\newblock Minimum $s$-$t$ cut of a planar undirected network in {$O(n\log^2
  n)$} time.
\newblock \emph{SIAM J. Comput.} 12:71--81, 1983.

\bibitem{r-mct-74}
G.~Ringel.
\newblock \emph{Map Color Theorem}.
\newblock Springer-Verlag, 1974.

\bibitem{ry-shmcp-68}
G.~Ringel and J.~W.~T. Youngs.
\newblock Solution of the {Heawood} map-coloring problem.
\newblock \emph{Proc. Nat. Acad. Sci. USA} 60:438--445, 1968.

\bibitem{s-cape-03}
A.~Schrijver.
\newblock \emph{Combinatorial Optimization: Polyhedra and Efficiency}.
\newblock Algorithms and Combinatorics~24. Springer-Verlag, 2003.

\bibitem{s-hco-05}
A.~Schrijver.
\newblock On the history of combinatorial optimization (till 1960).
\newblock \emph{Handbook of Discrete Optimization}, 1--68, 2005. Elsevier.

\bibitem{st-dsdt-83}
D.~D. Sleator and R.~E. Tarjan.
\newblock A data structure for dynamic trees.
\newblock \emph{J. Comput. Syst. Sci.} 26(3):362--391, 1983.

\bibitem{s-ctcgt-93}
J.~Stillwell.
\newblock \emph{Classical Topology and Combinatorial Group Theory}, 2nd
  edition.
\newblock Graduate Texts in Mathematics~72. Springer-Verlag, 1993.

\bibitem{tm-spltm-09}
S.~Tazari and M.~M{\"u}ller-Hannemann.
\newblock Shortest paths in linear time on minor-closed graph classes, with an
  application to {Steiner} tree approximation.
\newblock \emph{Discrete Appl. Math.} 157:673--684, 2009.

\bibitem{t-egnsn-90}
C.~Thomassen.
\newblock Embeddings of graphs with no short noncontractible cycles.
\newblock \emph{J. Comb. Theory Ser. B} 48(2):155--177, 1990.

\bibitem{v-slpfm-89}
P.~M. Vaidya.
\newblock Speeding-up linear programming using fast matrix multiplication.
\newblock \emph{Proc. 30th IEEE Symp. Found. Comput. Sci.}, 332--337, 1989.

\bibitem{w-mstfp-97}
K.~Weihe.
\newblock Maximum $(s, t)$-flows in planar networks in {$O(|V|\log|V|)$}-time.
\newblock \emph{J. Comput. Syst. Sci.} 55(3):454--476, 1997.

\bibitem{z-tc-05}
A.~Zomorodian.
\newblock \emph{Topology and Computing}.
\newblock Cambridge University Press, 2005.

\bibitem{zc-lh-07}
A.~Zomorodian and G.~Carlsson.
\newblock Localized homology.
\newblock \emph{Proc. IEEE Int Conf. Shape Modeling Appl. (SMI)}, 189--198,
  2007.

\end{thebibliography}
